### 2021-03-10 authors.tex

%------ affiliation shortcuts
\newcommand{\SPAno}{1}
\newcommand{\OzGravMonashno}{2}
\newcommand{\CITno}{3}
\newcommand{\MITno}{4}
\newcommand{\Kavlino}{5}

% %--------------------------
% \newcommand{\SPA}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
% \newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
% \newcommand{\MIT}{LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
% \newcommand{\Kavli}{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, \\ 77 Massachusetts Ave, Cambridge, MA 02139, USA}
% \newcommand{\CIT}{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}

% \author{
% \parbox{\textwidth}{
% A.~Vajpeyi$^{\SPAno,\OzGravMonashno}$, %ORCID:0000-0002-4146-1132
% R.~Smith$^{\SPAno,\OzGravMonashno}$, %ORCID:0000-0001-8516-3324
% E.~Thrane$^{\SPAno,\OzGravMonashno}$, %0000-0002-4418-3895
% G.~Ashton$^{\SPAno,\OzGravMonashno}$, %ORCID:0000-0001-7288-2231 
% T.~Alford$^\CITno$, 
% S.~Garza$^\CITno$,
% M.~Isi$^{\MITno,\Kavlino}$, %0000-0001-8040-9807
% J.~Kanner$^\CITno$, 
% T.J.~Massinger$^\CITno$,
% L.~Xiao$^\CITno$ %0000-0003-2703-449X
% }\vspace{0.2cm}\\
% $^1$\SPA\\
% $^2$\OzGravMonash\\
% $^3$\CIT\\
% $^4$\MIT\\
% $^5$\Kavli\\
% }

\author[\SPAno,\OzGravMonashno]{A.~Vajpeyi}
\author[\SPAno,\OzGravMonashno]{R.~Smith}
\author[\SPAno,\OzGravMonashno]{E.~Thrane}
\author[\SPAno,\OzGravMonashno]{G.~Ashton}
\author[\CITno]{T.~Alford}
\author[\CITno]{S.~Garza}
\author[\MITno, \Kavlino]{M.~Isi}
\author[\CITno]{J.~Kanner}
\author[\CITno]{T.J.~Massinger}
\author[\CITno]{L.~Xiao}

\affil[\SPAno]{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
\affil[\OzGravMonashno]{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
\affil[\MITno]{LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
\affil[\Kavlino]{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, \\ 77 Massachusetts Ave, Cambridge, MA 02139, USA}
\affil[\CITno]{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}