### 2021-04-15 authors.tex
edits for v2

%------ affiliation shortcuts
\newcommand{\SPAno}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
\newcommand{\OzGravMonashno}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
\newcommand{\CITno}{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
\newcommand{\MITno}{LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
\newcommand{\Kavlino}{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, 77 Massachusetts Ave, Cambridge, MA 02139, USA}

\author{A.~Vajpeyi}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{R.~Smith}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{E.~Thrane}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{G.~Ashton}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{T.~Alford}
\affiliation{\CITno}
\author{S.~Garza}
\affiliation{\CITno}
\author{M.~Isi}
\affiliation{\MITno}
\affiliation{\Kavlino}
\author{J.~Kanner}
\affiliation{\CITno}
\author{T.J.~Massinger}
\affiliation{\CITno}
\author{L.~Xiao}
\affiliation{\CITno}

% \affiliation[\OzGravMonashno]{}
% \affiliation[\MITno]}
% \affiliation[\Kavlino]{}
% \affiliation[\CITno]{}