### 2021-05-28 authors.tex
add todos

%------ affiliation shortcuts
\newcommand{\SPAno}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
\newcommand{\OzGravMonashno}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
\newcommand{\CITno}{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
\newcommand{\MITno}{LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
\newcommand{\Kavlino}{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, 77 Massachusetts Ave, Cambridge, MA 02139, USA}

\author{Avi~Vajpeyi}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{Rory~Smith}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{Eric~Thrane}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{Gregory~Ashton}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{Thomas~Alford}
\affiliation{\CITno}
\author{Sierra~Garza}
\affiliation{\CITno}
\author{Maximiliano~Isi}
\affiliation{\MITno}
\affiliation{\Kavlino}
\author{Jonah~Kanner}
\affiliation{\CITno}
\author{T. J.~Massinger}
\affiliation{\CITno}
\author{Liting~Xiao}
\affiliation{\CITno}

% \affiliation[\OzGravMonashno]{}
% \affiliation[\MITno]}
% \affiliation[\Kavlino]{}
% \affiliation[\CITno]{}