### 2021-08-04 authors.tex
Pre-MNRAS conversion

%------ affiliation shortcuts
\newcommand{\SPAno}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
\newcommand{\OzGravMonashno}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
\newcommand{\CITno}{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
\newcommand{\MITno}{LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
\newcommand{\Kavlino}{Department of Physics and Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, 77 Massachusetts Ave, Cambridge, MA 02139, USA}
\newcommand{\Royal}{Department of Physics, Royal Holloway, University of London, TW20 0EX, United Kingdom}

\author{Avi~Vajpeyi} % 0000-0002-4146-1132
\email{avi.vajpeyi@monash.edu}
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{Rory~Smith} %0000-0001-8516-3324
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{Eric~Thrane} %0000-0002-4418-3895
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\author{Gregory~Ashton} %0000-0001-7288-2231 
\affiliation{\SPAno}
\affiliation{\OzGravMonashno}
\affiliation{\Royal}
\author{Thomas~Alford} % 0000-0003-1942-1334
\affiliation{\CITno}
\author{Sierra~Garza}
\affiliation{\CITno}
\author{Maximiliano~Isi} %0000-0001-8830-8672
\affiliation{\MITno}
\affiliation{\Kavlino}
\author{Jonah~Kanner} % 0000-0001-8115-0577
\affiliation{\CITno}
\author{T. J.~Massinger} %0000-0002-3429-5025
\affiliation{\CITno}
\author{Liting~Xiao} %0000-0003-2703-449X
\affiliation{\CITno}

% \affiliation[\OzGravMonashno]{}
% \affiliation[\MITno]}
% \affiliation[\Kavlino]{}
% \affiliation[\CITno]{}