### 2022-01-21 referee_response/review_response.tex
Added ref review

% ADAPTED FROM https://github.com/hollina/template-referee-response

\documentclass[11pt,leqno]{article}

%------------------------------------------%
%                Packages
%------------------------------------------%

\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{microtype,xparse,tcolorbox}
\usepackage{graphicx}
\usepackage{natbib}	%Bibliography package. It handles citations and can easily change formats. This is done at the end of the document
\usepackage[utf8]{inputenx}% For proper input encoding
\usepackage{adjustbox} %Handles resizing of tables, figures, etc. Based off of page and/or line width/height 
% Packages for tables
\usepackage{booktabs}% For Pretty tables
\usepackage{threeparttable}% For Notes below table
\usepackage{rotating}% To Rotate Table
\usepackage{amsmath, amssymb,mathrsfs} %For using math
\usepackage{bm}
\usepackage{caption} 
\usepackage[list=true]{subcaption} %For having multiple figures within the same one. Figure 1, with part (a) and (b)
%\setcounter{lofdepth}{2}
\usepackage{setspace} %Single, Double Space, etc 
\usepackage[paperwidth=8.5in, paperheight=11in,margin=1in]{geometry} %For controlling the dimensions. Very useful for Posters, etc 
\usepackage{chngpage} 
\usepackage{everypage}%For page numbers on rotated pages
\usepackage[capposition=top]{floatrow}
\usepackage{accents}
\usepackage{float}
\usepackage{comment}
\usepackage{morefloats}
\usepackage{placeins}% To Create Float Barriers so the tables will stay in their sections
\usepackage{pdflscape}

\usepackage{siunitx} %This aligns tables by their decimal and handles processing of the numbers within the tables
  \sisetup{
    detect-mode,
    group-digits      = false,
    input-symbols     = {( ) [ ] - +},
    table-align-text-post = false,
    input-signs             = ,
    %parse-numbers=false,
    %scientific-notation = true,
        %round-mode              = places,
        %round-precision         = 2,
        %input-ignore={,},
    %input-decimal-markers={.},
    %group-separator={,},
        } 
\usepackage{grffile}
\usepackage{soul}
\usepackage{color}
\usepackage{caption}
%\captionsetup[figure]{justification=raggedright,singlelinecheck=off}
\sisetup{separate-uncertainty=true}
\usepackage{mathptmx}
\DeclareCaptionLabelFormat{blank}{}

%------------------------------------------%
%              Custom Commands
%------------------------------------------%
\newcommand\iid{i.i.d.} %Makes iid a command
\newcommand\pN{\mathcal{N}} %Makes the natural numbers a command

\setcounter{MaxMatrixCols}{20}

\newtheorem{theorem}{Theorem}%[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{assumption}[theorem]{Assumption}
\newenvironment{proof}{\paragraph{Proof:}}{\hfill$\square$ \\} %This fixes the problem with dynamic tables for some reason

%For Page Numbers while rotated
\newlength{\hfoot}
\newlength{\vfoot}
\AddEverypageHook{\ifdim\textwidth=\linewidth\relax
\else\setlength{\hfoot}{-\topmargin}%
\addtolength{\hfoot}{-\headheight}%
\addtolength{\hfoot}{-\headsep}%
\addtolength{\hfoot}{-.5\linewidth}%
\ifodd\value{page}\setlength{\vfoot}{\oddsidemargin}%
\else\setlength{\vfoot}{\evensidemargin}\fi%
\addtolength{\vfoot}{\textheight}%
\addtolength{\vfoot}{\footskip}%
\raisebox{\hfoot}[0pt][0pt]{\rlap{\hspace{\vfoot}\rotatebox[origin=cB]{90}{\thepage}}}\fi}
%http://tex.stackexchange.com/questions/209685/landscape-mode-and-page-numbering

%------------------------------------------%
%      Title Page Info
%------------------------------------------%
\title{Replies for MN-21-1316:\\
 ``A search for intermediate mass black holes mergers in the second LIGO-Virgo observing run with the Bayes Coherence Ratio
''}
\author{AUTHOR}
\date{\today}
\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

\begin{document}

\onehalfspacing

\section*{Response to the referee}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Thank you for your helpful comments and guidance. We describe how we have addressed all of your concerns below. \\
\\
Your comments are below,  \fcolorbox{black}{black!10}{{\ttfamily boxed in with a gray background and different font}}. Our responses are in plain text. Any included text from the new manuscript is \fbox{boxed in with a white background and the} \\
\fbox{same font as this text} with the page number noted at the top of the quoted text. 
Any added comments or other differences from the quoted manuscript text and the text appearing in this report {\color{red} {\ttfamily[appears in brackets with a different font and a red color]}}. \\
\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Referee: Comment 1} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{tcolorbox}[left = 1em, top = 1ex, bottom = 1ex, colupper=black, colback=black!10, adjusted title = Referee: Comment 1]
    \setlength\parindent{2em}
	\noindent
	\ttfamily

	Title: black holes mergers -> black hole mergers
\end{tcolorbox}

We have adjusted the title.

\end{document}