### 2022-01-21 referee_response/example.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Referee 1: Short description of comment
\subsection*{Referee: Comment 1} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Exact comment- copied and pasted
\begin{tcolorbox}[left = 1em, top = 1ex, bottom = 1ex, colupper=black, colback=black!10, adjusted title = Referee: Comment 1]
    \setlength\parindent{2em}
	\noindent
	\ttfamily

	WORD FOR WORD COPY AND PASTE OF COMMMENT
\end{tcolorbox}

% Response

\paragraph{Short indicator of what the point is (e.g., Event study coefficients are hard to read)} Answer

\paragraph{Short indicator of what the point is (e.g., Event study should use -1 as reference category)} Answer

In response we have added the following text to the manuscript (beginning on \hl{page XX}).

% Text added to manuscript    
	\begin{tcolorbox}[left = 1em, top = 1ex, bottom = 1ex, colupper=black, colback=white, adjusted title = From page \hl{XX}]

       EXACT TEXT FROM MANUSCRIPT

        \begin{align}
            y_{it} = &~\beta X + \epsilon_{it} \notag
        \end{align}

        We report all coefficient estimates in Table~2. {\color{red} {\ttfamily[Table~\ref{tab:main_results}in this report.}} \\

        {\color{red} {\ttfamily[text continued on next page]}} \\
    \end{tcolorbox}