### 2022-01-21 titlepage_mnras.tex

%------ affiliation shortcuts
\newcommand{\SPA}{School of Physics and Astronomy, Monash University, Clayton VIC 3800, Australia}
\newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
\newcommand{\CIT}{LIGO Laboratory, California Institute of Technology, Pasadena, CA 91125, USA}
\newcommand{\MIT}{LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, MA 02139, USA}
\newcommand{\Kavli}{Department of Physics and Kavli Institute for Astrophysics and Space Research,\\ Massachusetts Institute of Technology, 77 Massachusetts Ave, Cambridge, MA 02139, USA}
\newcommand{\Royal}{Department of Physics, Royal Holloway, University of London, TW20 0EX, United Kingdom}

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% % Title of the paper, and the short title which is used in the headers.
% % Keep the title short and informative.
% \title[MNRAS \LaTeX\ guide for authors]{\textit{Monthly Notices of the Royal Astronomical
%   Society}:  \LaTeX\ guide for authors}
\title[An IMBH search in O2 using the BCR]{A search for intermediate-mass black holes mergers in the \\second LIGO--Virgo observing run with the Bayes Coherence Ratio}

% % The list of authors, and the short list which is used in the headers.
% % If you need two or more lines of authors, add an extra line using \newauthor
% \author[K. T. Smith]{Keith T. Smith$^{1}$%
% \thanks{Contact e-mail: \href{mailto:mn@ras.ac.uk}{mn@ras.ac.uk}}%
% \thanks{Present address: Science magazine, AAAS Science International, \mbox{82-88}~Hills Road, Cambridge CB2~1LQ, UK}%
% \\
% % List of institutions
% $^{1}$Royal Astronomical Society, Burlington House, Piccadilly, London W1J 0BQ, UK}

\author[Vajpeyi et al.]{Avi~Vajpeyi$^{1,2}$\thanks{avi.vajpeyi@monash.edu},
    Rory~Smith$^{1,2}$,
    Eric~Thrane$^{1,2}$,
    Gregory~Ashton$^{1,2,3}$,
    Thomas~Alford$^{4}$,
    \newauthor
    Sierra~Garza$^{4}$,
    Maximiliano~Isi$^{5,6}$,
    Jonah~Kanner$^{4}$,
    T. J.~Massinger$^{4}$,
    Liting~Xiao$^{4}$
\\
$^{1}$ \SPA \\
$^{2}$ \OzGravMonash \\
$^{3}$ \Royal \\
$^{4}$ \CIT \\
$^{5}$ \MIT \\
$^{6}$ \Kavli 
}

% % These dates will be filled out by the publisher
\date{Last updated XXX; in original form XXX}

% % Enter the current year, for the copyright statements etc.
\pubyear{2021}